---
title: "Carga de DF"
author: "Curso de Estadística Descriptiva"
date: "23/12/2018"
output: html_document
---

#probando la carga

```{r}
df2 = read.table("../../data/StudentsData",
                 header = TRUE,
                 sep="")
head(df2,10)
              
```
```{r}
flags = read.table("../../data/flag.data",sep=",",stringsAsFactors = )
head(flags
     )

```


# Carga de ficheros local

Breed (1/5/8), 
Sale Price, 
Yearling height at shoulder (in.), 
Fat Free Body (lbs.), 
Percent Fat-free body, 
Frame -- scale from 1 (small) to 8 (large), 
Back fat (in.), 
sale height at shoulder (in.) 
sale weight (lbs.) of three breeds of bulls.


```{r}
df = read.table("..\..\data\bulls.dat,
                header = FALSE, 
                col.names = c("breed", "sale_price", "shoulder",
                          "fat_free", "percent_ff", "frame_scale",
                          "back_fat", "sale_height", "sale_weight"),
                sep = "", dec = ".")
head(df)




# Carga desde URL

```{r}
df2 = read.table("https://maitra.public.iastate.edu/stat501/datasets/bulls.dat",
                 header = FALSE, 
                col.names = c("breed", "sale_price", "shoulder",
                          "fat_free", "percent_ff", "frame_scale",
                          "back_fat", "sale_height", "sale_weight"),
                sep = "", dec = ".")
head(df2)
str(df2)
```


# Factores en un DF

```{r}
letter = read.table("https://people.sc.fsu.edu/~jburkardt/data/csv/letter_frequency.csv",
             stringsAsFactors = TRUE,
            header = TRUE,
            sep = ",")
letter
```

# Guardar un DF

```{r}
write.table(letter, file = "../../data/frecuencia_letras.txt", dec = ".")
frecuencia = read.table("../../data/frecuencia_letras.txt", header = TRUE, dec = ".")
head(frecuencia1)
```


# Crear un DF

```{r}
gender = c("H", "M", "M", "M", "H")
age =    c( 23,  45,  20,  30,  18)
family = c(  2,   3,   4,   2,   5)
df5 = data.frame(genero = gender, edad = age, familia = family, stringsAsFactors = TRUE)
row.names(df5) = c("P1", "P2", "P3", "P4", "P5")
df5
str(df5)
dimnames(df5) = list(
  c("Antonio", "Ricardo", "JuanGabriel", "María", "Margarita"),
  c("Sexo", "Años", "MiembrosFamilia")
)
df5
df5 = rbind(df5, c("H", 30, 1))
df5
df5$Sexo = as.character(df5$Sexo)
df5$Ingresos = c(10000, 12000, 15000, 12000, 20000, 10000)
```


```{r}
gender = c("H", "M", "M", "M", "H")
age =    c( 23,  45,  20,  30,  18)
family = c(  2,   3,   4,   2,   5)
df5 = data.frame(genero = gender, edad = age, familia = family, stringsAsFactors = TRUE)
df5[df5$genero=="M", ] -> df_m
str(df_m)
df_m = droplevels(df_m)
str(df_m)
```

# Tidyverse
```{r}
library(tidyverse)
iris_petal = select(iris, starts_with("Petal"))
head(iris_petal)
iris_length = select(iris, ends_with("Length"))
head(iris_length)
```

# Subset
```{r}
subset(iris, Species == "versicolor", select = c(1,3)) -> versicolor
rownames(versicolor) = 1:nrow(versicolor)
head(versicolor, 5)
str(versicolor)
```

